---
title: "Nina graphics in figures"
author: "Jens Åström"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nina colors and logo in R-figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=F, message=F}
require(NinaR)
```


Intro
==============
The `NinaR` package provides functions for using the recommended NINA colors in R-figures. Currently, the package provides palette functions for providing accent colors and the colors found in the logotype, and a function that adds transparency to colors as an alpha channel.

In addition, the package provides a function for adding the Nina logo to a plot.

NinaPalette and NinaLogoPalette
===========

R uses a default palette made up of the colors `r palette()`. These colors can be selected in figures by numbers e.g. `col=2` means the color red. The function `NinaPalette` and `NinaLogoPalette` provides similar functionality by providing a new palette in line with the NINA graphical handbook. 

To replace the default palette with the Nina palette:

```{r palette}
palette(NinaPalette())
palette()
```

Alternatively, the palette can be used directly in functions:

```{r pallettefunction, fig.width=7, fig.height=5}
require(NinaR)
plot(10:1, cex=2, pch=16, col=NinaPalette())
```

Transparency can be added with the `addAlpha` function.

```{r addAlpha, fig.width=7, fig.height=5}
set.seed(12345)
barplot(runif(5), col=addAlpha(NinaPalette(), 0.4))
```


AddLogo
=============

The function `addLogo`, not surprisingly, adds a Nina logo to a figure. The default behaviour adds the logo to the right bottom corner, but the position and size can be specified to suit most needs.

```{r addlogo, fig.width=7, fig.height=5}
plot((1:10)^2, 1:10, col=NinaPalette(), cex=4, pch=16, las=1)
addLogo()
```

The logo can also be added in the background, with a faded tint, with help from the package `grImport` (loaded automatically). Note that you have to initiate a plot before adding the logo, and therefore you have to elements on top of the logo manually afterwards.

```{r addbackgroundlogo, fig.width=7, fig.height=5}
plot((1:10)^2, 1:10, col=NinaPalette(), cex=4, pch=16, las=1, type="n")
addLogo(x = 0.5, y = 0.5, size = 1)
grid.rect(gp = gpar(fill = rgb(1, 1, 1, .6)))
points((1:10)^2, 1:10, col=NinaPalette(), cex=4, pch=16)
```


