---
title: "Nina grafik i figurer"
author: "Jens Åström"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nina colors and logo in R-figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Intro
==============
The `NinaR` package provides functions for using the recommended NINA colors in R-figures. Currently, the package provides palette functions for providing accent colors and the colors found in the logotype, and a function that adds transparency to colors as an alpha channel.

In addition, the package provides a function for adding the Nina logo to a plot.

NinaPalette and NinaLogoPalette
===========

R uses a default palette made up of the colors `r palette()`. These colors can be selected in figures by numbers e.g. `col=2` means the color red. The function `NinaPalette` and `NinaLogoPalette` provides similar functionality. 

To replace the default palette with the Nina palette:

```{r}
palette(NinaPalette())
palette()
```

Alternatively, the palette can be used directly in functions:

```{r}
require(NinaR)
plot(10:1, cex=2, pch=16, col=NinaPalette())
```
