---
title: "Nina graphics in figures"
author: "Jens Åström"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nina colors and logo in R-figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=F, message=F}
require(NinaR)
require(grid)
require(ggplot2)
knitr::opts_chunk$set(
  dev = 'svg'
)
options(rmarkdown.html_vignette.check_title = FALSE)
```


Intro
==============
The `NinaR` package provides functions for using the recommended NINA colors in R-figures. Currently, the package provides a standard palette function for providing accent colors and the colors found in the logotype, and a function that adds transparency to colors as an alpha channel. It also contains a ggplot2 color scale for fills and colors.

In addition, the package provides a function for adding the Nina logo to a plot.

ninaPalette and ninaLogoPalette
===========

R uses a default palette made up of the colors `r palette()`. These colors can be selected in figures by numbers e.g. `col=2` means the color red. The function `NinaPalette` and `NinaLogoPalette` provides similar functionality by providing a new palette in line with the NINA graphical handbook. 

To replace the default palette with the Nina palette:

```{r palette}
palette(ninaPalette())
palette()
```

Alternatively, the palette can be used directly in functions:

```{r pallettefunction, fig.width=7, fig.height=5}
require(NinaR)
plot(10:1, cex=2, pch=16, col=ninaLogoPalette())
```

Transparency can be added with the `addAlpha` function.

```{r addAlpha, fig.width=7, fig.height=5}
set.seed(12345)
barplot(runif(5), col=addAlpha(ninaPalette(), 0.4))
```


addLogo
=============

The function `addLogo`, not surprisingly, adds a Nina logo to a figure. The default behaviour adds the logo to the right bottom corner, but the position and size can be specified to suit most needs.

```{r addlogo, fig.width=7, fig.height=5}
plot((1:10)^2, 1:10, col=ninaPalette(), cex=4, pch=16, las=1)
addLogo()
```

The logo can also be added in the background, with a faded tint, with help from the package `grImport` (loaded automatically). Note that you have to initiate a plot before adding the logo, and therefore you have to elements on top of the logo manually afterwards.

```{r addbackgroundlogo, fig.width=7, fig.height=5}
plot((1:10)^2, 1:10, col=ninaPalette(), cex=4, pch=16, las=1, type="n")
addLogo(x = 0.5, y = 0.5, size = 1)
grid.rect(gp = gpar(fill = rgb(1, 1, 1, .6)))
points((1:10)^2, 1:10, col=ninaPalette(), cex=4, pch=16)
```

ggplot2
============

I have also added two ggplot scales, `scale_fill_nina` and `scale_color_nina`. As the names suggests, they should be used in cases where you define colors by the fill or color aesthetics, respectively. Both have the options to define the palettes, which are "main" and "logo" at the moment, and if the scale is discrete or not (continuous). These aren't the prettiest examples, but you get the picture. I plan to add more palettes later on, for example some to use with continous gradients.

```{r scale_fill_nina_continous, fig.width=7, fig.height=5}
g <- ggplot(faithfuld, aes(waiting, eruptions))
g + geom_raster(aes(fill = density)) +
  scale_fill_nina(discrete = F)
```

```{r scale_fill_nina_discrete, fig.width=7, fig.height=5}
g <- ggplot(mpg, aes(class))
g + geom_bar(aes(fill = drv)) +
   scale_fill_nina(palette = "logo")
```

```{r scale_color_nina_discrete, fig.width=7, fig.height=5}
g <- ggplot(mpg, aes(cyl, hwy)) 
g + geom_point(aes(colour = drv)) +
   scale_color_nina(discrete = T)
```

```{r scale_color_nina_continuous, fig.width=7, fig.height=5}
g <- ggplot(mpg, aes(cty, hwy)) 
g + geom_point(aes(colour = cyl)) +
   scale_color_nina(discrete = F)
```

add_logo for ggplots
=============
There is also a ggplot version of the add logo funtion, called add_logo. This adds an svg version of the logo to an existing ggplot. It requires some tinkering with the position, size, and the thickness of the line strokes of the logo. As of yet it doesn't have a transparancy option.


```{r add_logo, fig.width=7, fig.height=5}
g <- ggplot(mpg, aes(cty, hwy))  +
  geom_point(aes(colour = cyl)) +
   scale_color_nina(discrete = F) +
  add_logo(xmin = 25,
           xmax = 30,
           stroke_scale = 0.3)

g

```



